---
title: "Statistical Inference Project-2"
author: "Adekunle Adeniran"
date: "2023-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PART 2 Basic Inferential Data Analysis

Loading the required libraries and take a peek at the dataset
```{r, echo=TRUE}
library(stats)
library(ggplot2)
library(kableExtra)
data("ToothGrowth")
head(ToothGrowth)
```
# Summary of Data
```{r, echo=TRUE}
summary(ToothGrowth)
```



```{r, echo=TRUE}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity",) +
    facet_grid(. ~ supp) +
    xlab("Dose in miligrams") +
    ylab("Tooth length") +
    guides(fill=guide_legend(title="Supplement type"))
```

# Using Confidence Intervals to compare tooth growth by supp and dose.

```{r, echo=TRUE}
unique(ToothGrowth$dose)
```

There are 3 dose groups: 0.5, 1 and 2

The Graph below shows the relationship between Tooth length and Dose

```{r, echo=TRUE}
ggplot(aes(x=factor(dose), y=len), data=ToothGrowth) + geom_boxplot(aes(fill= factor(dose))) + ggtitle("Plot of length by dose") + xlab("Dose (mg)") + ylab("Tooth length") + labs(fill = "Dose (mg)")
```
*T-test for dose 0.5 mg*:

```{r, echo=TRUE}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
```


*T-test for dose 1 mg*:

```{r, echo=TRUE}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
```

*T-test for dose 2 mg*:
```{r, echo=TRUE}
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])
```

```{r, echo=TRUE}
dose <- c(0.5, 1.0, 2.0)
p_value <- c(0.0064, 0.0010, 0.9639)
conf.int <-c("1.72, 8.78","2.80, 9.06","-3.80, 3.64")
decision <- c("Reject null", "Reject null", "Do not reject null")

knitr::kable(data.frame(dose, conf.int, p_value, decision), align = "cccc")
```

As anticipated, the p-values for doses 0.5 and 1.0 are expected to be very low due to the substantial mean differences between them.

Consequently, for doses 0.5 and 1.0, since the p-values fall below 0.5, we can reject the null hypotheses asserting that there is no difference in tooth growth among the supplement types. However, for dose 2.0 mg/day, the null hypothesis can be retained as the p-value exceeds 0.5.


# Conclusion

The fundamental assumption underlying the results is that the sample is a representative depiction of the population, and the variables are independent and identically distributed (IID) random variables.

Regarding the t-test, two key assumptions are taken into account:

1. The data is not paired, indicating independence.
2. The variances are unequal.

Given these considerations, upon reviewing the t-test results, it is observed that supplement type OC proves to be more effective than VC for doses below 1.0. However, at a dose of 2.0 mg/day, there is no discernible difference between the supplement types.

